### 2023-01-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@0ac08f659b21331459213cd1ce9ba8400f5f6fb3 🚀

\hypertarget{general-background}{%
\subsection{General background}\label{general-background}}

Why is it useful to think about interactions as probabilistic events?

\begin{itemize}
\tightlist
\item
  A biological interaction is probabilistic since two taxa co-occurring
  does not mean they are going to meet (e.g., think of species relative
  abundances). Also, two individuals meeting does not mean that an
  interaction will occur (e.g., a lion crossing paths with a gazelle
  does not mean predation).
\item
  Species interactions are contextual on the environment and on the
  physiological state of both species (or individuals).
\item
  Representing trophic interactions as probabilistic events helps us
  predict food webs across time and space and assess their spatial
  variability.
\end{itemize}

\textbf{Papers:} Poisot \emph{et al.} (2016)

\hypertarget{problem-and-objectives}{%
\subsection{Problem and objectives}\label{problem-and-objectives}}

Why should we use probabilities with caution when working with food webs
and other ecological networks? What is the objective of this paper?

\begin{itemize}
\tightlist
\item
  There are different ways to define and interpret interaction
  probabilities depending on the study system and on the method used to
  build probabilistic food webs.
\item
  It is important to document and define what we mean by an interaction
  probability because different definitions can have different
  ecological and statistical implications and interpretation.
\item
  This paper aims to outline some of the challenges in interpreting
  interaction probabilities in food webs and propose an approach to
  thinking about them. Using clear and simple mathematical equations, we
  distinguish different meanings of probabilistic interactions.
\end{itemize}

@@

\textbf{Papers:} Strydom \emph{et al.} (2022)

@@

\textbf{Papers:} there might be something in these McLeod \emph{et al.}
(2020); McLeod \emph{et al.} (2021); Wood \emph{et al.} (2015)

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S.J., Gravel, D., Chu, C., Cirtwill, A.R., Fortin,
M.-J., \emph{et al.} (2021). Sampling and asymptotic network properties
of spatial multi-trophic networks. \emph{Oikos}, n/a.

\leavevmode\hypertarget{ref-McLeod2020EffSpe}{}%
McLeod, A.M., Leroux, S.J. \& Chu, C. (2020). Effects of species traits,
motif profiles, and environment on spatial variation in multi-trophic
antagonistic networks. \emph{Ecosphere}, 11, e03018.

\leavevmode\hypertarget{ref-Poisot2016StrProa}{}%
Poisot, T., Cirtwill, A.R., Cazelles, K., Gravel, D., Fortin, M.-J. \&
Stouffer, D.B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, 7, 303--312.

\leavevmode\hypertarget{ref-Strydom2022FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M.J., \emph{et al.} (2022). Food web reconstruction through phylogenetic
transfer of low-rank network representation. \emph{Methods in Ecology
and Evolution}, n/a.

\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
Wood, S.A., Russell, R., Hanson, D., Williams, R.J. \& Dunne, J.A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, 5, 3769--3782.

>>>>

Cataloging species interactions across space is a gargantuan task. At
the core of this challenge lies the spatiotemporal variability of
ecological networks (Poisot \emph{et al.} 2015), which makes documenting
the location and timing of interactions difficult. Indeed, it is not
sufficient to know if two species have the biological capacity to
interact to infer the realization of this interaction at a specific time
and space. Taking food webs as an example, a predator and its potential
prey must first co-occur on the same territory in order for a trophic
interaction to take place. They must then cross paths, which can happen
only if their relative abundances in the ecosystem is high enough and
their phenology matches. Finally, the interaction occurs only if the
predator has a desire to consume its prey and is able to capture and
ingest it. Environmental (e.g. temperature and presence of shelters) and
biological (e.g.~physiological state of both individuals and
availability of other prey species) factors contribute to this
variability by impacting species co-occurrence and the realization of
their interactions. In this context, it is unsurprising that
computational methods are being developed in ecology to help alleviate
the colossal sampling efforts required to document species interactions
across time and space (Strydom \emph{et al.} 2021).

The recognition of the variability of species interactions and the
emergence of numerical methods have led ecologists to rethink their
representation of ecological networks, slowly moving from a binary to a
probabilistic view of species interactions
(\textbf{Poisot2016StrProa?}). This has several benefits. For example,
probabilities represent the limit of our knowledge about species
interactions and can indicate the expected frequency of two species
interacting with each other. They are also very helpful in predictive
models when modeling the spatial distribution and temporal variability
of ecological networks, generating new ecological data, and identifying
priority sampling locations.

However, representing species interactions probabilistically can also be
challenging. Beyond methodological difficulties in estimating these
numbers, there are important conceptual challenges in defining what we
mean by ``probability of interactions.'' Because this ecological
representation is still in its infancy, there is no clear definition
found in the literature. In this contribution, we outline different ways
to define and interpret interactions probabilities in network ecology
and propose an approach to thinking about them. These definitions mostly
depend on the study system (e.g.~local network or metaweb) and on the
method used to generate them. We show that different definitions can
have different ecological implications, especially related to spatial,
temporal, and taxonomic scaling. Although we will focus on food webs,
our observations and advice can be applied to all types of ecological
networks, from plant-pollinator to host-parasite networks. Overall, we
argue that probabilities should be better documented, defined
mathematically, and used with caution when describing species
interactions.

@@

\textbf{Papers:} (\textbf{Strydom2022FooWeb?})

@@

\textbf{Papers:} there might be something in these McLeod \emph{et al.}
(2020); (\textbf{McLeod2021SamAsy?}); (\textbf{Wood2015EffSpa?})

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-McLeod2020EffSpe}{}%
McLeod, A.M., Leroux, S.J. \& Chu, C. (2020). Effects of species traits,
motif profiles, and environment on spatial variation in multi-trophic
antagonistic networks. \emph{Ecosphere}, 11, e03018.

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B-Biological
Sciences}, 376, 20210063.
